Adopting agile in the large: motivations, success factors and challenges

### 2014-05-07 journal/article.tex
Work in progress

@@

* Hard to create roadmap between teams --> Teams see only their own priorities

@@

>>>>

% The software produced in a large organization may have size and complexity that
% agile struggles to handle.

The primary studies revealed challenges on two areas which relate particularly
to large scale. First, a large software development endeavour must handle
integrating the work products of many teams. Second, a large organization will
have management challenges not seen on small scale.

%% ?? -->
% Include this ???
% Large organization put agile methods under the test of scaling up, and problems
% were encountered. One case reported use of agile across hundreds of teams, but
% management could not get a roll up of the progress without having strict rules
% on how status was reported [P48]. Schnitter and Mackert [P45] report that a
% maximum workable size in a project managed by Scrum was 130 people including 7
% development teams.

Many cases reported that introducing agile practices soon yielded improvements
on the team level, but as a part of a large transformation teams needed to
change their external behavior also.
The initial success was followed by challenges when the teams needed to work as
parts of the larger surrounding orgaization [P13, P17, P24]. Introducing agile
had created flexibility on a team level, but it could not be fully utilized as
the surrounding organization was not responsive enough [P26].
There were dependencies between teams in the agile model as well, and the
dependencies made managing development difficult [P9, P10].

In the agile model it was hard for managers to get an overview of the progress
of the release [P17]. The status of features was hard to tell because different
aspects of features were sepatated across teams [P17]. Smits and Rilliet [P48]
report that providing executive overview of the development programs was limited
by the tools used to manage agile projects.

Some organizations attempted to initially create a model where teams operate
autonomously, which is well in line with agile principles. However, a number of
issues arose from independence. For instance, teams had to balance between the
team's own goals and broader goals of the entire organization, and some teams
chose focus on their own goal [P7]. Coordination was obstructed by independent
teams that did not respect the larger context they were working in [P13].
Coordination was also difficult when different teams were working for different
customers but the applications being built were interdependent [P29]. One case
reported that allowing teams to have different sprint durations created delays
in delivery [P51].

Also technical problems relating to inter-team coordination were reported.
Integrating the products of teams was seen as a problem in some cases [P5, P50],
and lack of standardized build scripts was mentioned as a problem in one case
[P27].
There was also problems in synchronizing definitions of software interfaces
between teams, and dependencies in code caused problems in larger features
[P26].
One case reported that team centricity in the agile model created a fragile
architecture, divergence in coding style, and even distrust between teams [P24].

Many studies presented solutions for the inter-team challenges, which also
highlight the challenges themselves. The general model of solving inter-team
problems was introducing an additional layer of coordination, such as a
Scrum-of-Scrums team [P43], a similar product team [P45], or an entire
customized release framework [P13]. In one case integration problems were
controlled by prioritizing code quality and maintaining existing functionality
over new features [P13]. Finally, the choice of tools was reported to have an
impact on the coordination effort [P17].

[P43] Transformation is described. Focus on “scaling up” part of adoption.

% --

* First cross team dependencies were reduced --> Problems with too much dependencies
* Traditional dependency management limited flexibility, which was demanded --> Inflexibility, work intensive, adversarial relationships
[P34]
From the onset, we sought to reduce cross-team dependencies as much as possible,
allowing each of the Scrum teams to “run” at their fastest pace.  However, as
the number of concurrent work streams increased, so did the number of
dependencies.  In particular, we found that adopting a SOA ran counter to our
efforts.
We quickly faced the prospect of implementing and integrating with a sizeable
set of centralized services that would be consumed by multiple project teams.
Within a few months, we had accumulated a significant number of integration
points, preventing us from delivering fully production-ready software until they
were resolved.
Accepting the fact that these dependencies would be part of the project, we
sought to create a lightweight and efficient process for resolving dependencies
and integration points.  We recognized that traditional approaches to dependency
management tend to place significant constraints on the sequence of work,
reducing the teams’ ability to independently prioritize their backlogs. 
Furthermore, highly contractual, document-centric approaches to defining
integration points are work-intensive, prone to error and can create adversarial
relationships between teams.  We wanted to establish an approach based on direct
communication and collaboration.
Our initial integration efforts involving five of the module teams of the Core
Application were reasonably successful. After some initial scrambling, the teams
self-organized and established a set of procedures for managing inter-module
integration.

* Worked decently with 5 teams and SoS --> Problems with global deployment
[P34]
However, our efforts to facilitate cross-team collaboration within the Core
Application benefited heavily from the fact that the five module teams were all
closely co-located, within the same organizational structure and employed an
active “Scrum of Scrums” model to ensure communication and coordination between
the teams.  As we pushed this approach out further across the program and the
globe, we began to experience significant challenges.

* Problem: Agile teams were thinking inside the team room walls
* Success: Thinking of the teams that will integrate to your work
* Failure: Thinking only inside the team walls
[P34]
We found that we had varying degrees of success in managing heavily-used
services.  In those cases where we were successful, it was invariably due to a
small group of people from both the consuming and providing teams taking
ownership of the services and proactively seeking out future integration points
and potential challenges.  Conversely, in those situations where such ownership
did not exist and people remained within their team walls, we invariably
experienced integration challenges: services consistently failing to meet the
consumer’s requirements and integration points that were poorly tested and
failed in system testing.

* Global effort problems --> Communication regressed to contractual --> Narrow communication channels were used: documents, e-mail
* Success: Individuals that saw effort in wide-channel communication, Individuals that thought outside team room walls
[P34]
As more and more integration occurred across the globe, we witnessed a tendency
to revert back to contractual forms of communication. As always, distributing
the development team significantly increases the challenge to directly
communicate and collaborate.  The natural tendency to rely on e-mail due to
differences in time zones frequently led to communication breakdowns and
unnecessary blockages – forming barriers to feedback. In some instances,
individuals broke through these barriers using instant messaging products, web
cams and adjusted working hours to facilitate as much direct communication as
possible.  At the same time other teams regressed, relying on documentation and
establishing an almost adversarial relationship with their integration partners.

* Large program problems --> Green builds hard to maintain
* Tools insufficient for large
[P34]
Given the size and distribution of our development team and the fact that the
program produced a single, integrated enterprise application, the complexity of
the build process and the level of effort to maintain it was significant.   As
the program grew in size, we recognized that the tool set with which we began
the project was no longer sufficient.  We began experiencing nearly continuous
broken builds due to conflation and timing issues on the build server.  At this
point, we began a lengthy initiative to replace the toolsets and revamp the
entire build process using AntHillPro and Subversion.  As a result, we created a
build process that effectively managed our myriad of dependencies and supported
a total build of the application in a reasonable time.

* Problems: Integration and dependency management
* Goal: Foster communication and collaboration instead of rigid processes
* Problems: Team avoiding external distractions, balance between team membership and project membership --> CI and green builds did not work
* Problems: Technical leads must work outside of team --> To coordinate
* Solution: Balance between own goals and project goals --> Think outside team room
* The team must think of others than them selves, too
[P34]
Over the lifetime of the project we encountered two key challenges due to the
scope and scale of this effort: integration and dependency management and
continuous integration. When addressing these challenges, we sought to adhere to
the basic principals of the Agile manifesto as much as possible: making the
delivery of working software our overriding goal and fostering communication and
collaboration instead of dictating rigid processes. We sought to empower
individuals and teams to make decisions and formulate their own processes, and
to take calculated risks without fear of reprisals.

However, we found that some of the qualities and practices that made the teams
work effectively in isolation actually reduced their effectiveness in the
large-scale environment.
For example: in smaller projects, we work aggressively to avoid external
distractions and encourage team members to focus exclusively on their individual
team goals as expressed in the backlog.  In large-scale projects, team members
must participate in cross-team activities and allocate time to project-wide
activities.  In particular, team technical leads were required to dedicate
significant portions of their time outside of the team room.  We found that many
of our team leads struggled with this task.  While they recognized the need for
cross-team communication, they resented context switching and the impact on
their time dedicated to the team’s sprint goals.
Furthermore, all of the teams struggled with balancing their identity as a team
member against their membership in the larger project. This was reflected most
keenly in our failure to sustain continuous integration and produce frequent
green builds.

Agile team must make trade-offs between stabilizing an application and
completing more story cards, the size of this project required teams to put
significantly more time into tracking down issues that were caused by other
teams or dependencies.   Many team members found this frustrating and
demoralizing.

@@

@@

* Transformation and scaling up are connected concepts. Although this research
  focuses on transformation process only, it can not be completely separated
  from discussion on scaling up, or on large scale agile implementation in
  general.

* Transformation of an organization to use agile can be seen as transforming
  also the agile method to suit the organization.
  --> A large scale agile model may require tailoring and trial over time

* It can be argued that any development effort with hundreds of teams will be
  difficult to coordinate

* Even though most of the primary studies are experience reports and subject to
  bias, this study can draw valuable conclusions. --> If a matter is being
  talked about it must be of some significance.
